---
title: "Example"
author: 'Emma Cliffe'
date: '2022'
documentclass: article
fontsize: 12pt
#documentclass: extarticle
#fontsize: 17pt
classoption: a4paper
geometry: margin=2.5cm
linkcolor: blue
output:
  word_document:
    toc: yes
  html_document:
    dev: svg
    fig_caption: yes
    keep_md: yes
    self_contained: no
    toc: yes
  pdf_document:
    dev: pdf
    toc: yes
    latex_engine: pdflatex
    fig_caption: true
    keep_tex: yes
header-includes: #\usepackage{helvet}\renewcommand{\familydefault}{\sfdefault}\setlength{\parindent}{0pt}\setlength{\parskip}{6pt}\renewcommand{\baselinestretch}{1.25}
---
<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  CommonHTML: { linebreaks: { automatic: true } },
  "HTML-CSS": { linebreaks: { automatic: true } },
         SVG: { linebreaks: { automatic: true } }
});
</script>

# Alternative formats

It is recommended that you [read this resource on the web](./arclengthInR.html) but you can also download [this resource as a PDF](./arclengthInR.pdf) or download [this resource as a Word document](./arclengthInR.docx). 

# Arc length 
We often need to know the length of a curve between two points,
e.g. what is the length of the ropes holding Clifton suspension bridge (see Exercise Sheet 3).

## Visualisation

Given a curve $y=y(x)$

\centering

```{r echo=FALSE, out.height="20%", screenshot.alt="arclength1.png"}
knitr::include_url("https://www.desmos.com/calculator/fd09vdozmt?embed")
```

\flushleft

Accessible interactive graph at https://www.desmos.com/calculator/t8dz6vlmnz


Let $S$ be the arc length and  $\Delta S$ a short section of it.

\centering

```{r echo=FALSE, out.height="20%", screenshot.alt="arclengthdx.png"}
knitr::include_url("https://www.desmos.com/calculator/qvrxafeubg?embed")
```

\flushleft

Accessible interactive graph at https://www.desmos.com/calculator/g5duc4kmfp



## Derivation of Arc Length

By Pythagoras' Theorem,
$$
\Delta S^2 \approx \Delta x^2+\Delta y^2
\qquad \Rightarrow\qquad
\left(\dfrac{\Delta S}{\Delta x}\right)^2 \approx 1+\left(\dfrac{\Delta y}{\Delta x}\right)^2
$$
As $\Delta x\to0$ this becomes an identity
$$
\left(\dfrac{dS}{dx}\right)^2 = 1+\left(\dfrac{dy}{dx}\right)^2
\qquad\Rightarrow\qquad
\dfrac{dS}{dx} = \sqrt{1+\left(\dfrac{dy}{dx}\right)^2}
$$
The arclength between $x=a$ and $x=b$ is then
$$
\begin{aligned}
  S(a,b) &= \int_a^b\dfrac{dS}{dx}dx\\
  &= \int_a^b\sqrt{1+\left(\dfrac{dy}{dx}\right)^2}dx.
\end{aligned}
$$

